### 2011-06-22 thesis/novelty-intro.tex
Added comments on using a constant P(novel)

%\subsection{Novelty Detection on a Variable Set of Features}
Note also that often novelty detection is applied on a fixed set of features
together with an assumption of a uniform unconditional probability.
On those cases $P(x)$ becomes a constant and therefore a novelty threshold
can be directly applied on $P(x|\overline{novel})$ as is the case in \cite{bishop1994novelty}.
But in the case where the set of features $x$ is variable it cannot be
discarded. There $P(x)$ also plays a role in levering all the conditional
probabilities on different sets of variables into the same measure units
such that a threshold can be implemented.

\section{Novelty Detection on the Conceptual Map}

@@

\subsection{Approximating the Conditional Probability}
The actual semantic mapping of the environment happens using the \emph{conceptual map}
introduced on \autoref{sec:conceptual-map}. During it the agent instantiates a
\emph{chain graph} to model distribution of the sensed variables according to the known
concepts and categories introduced as hidden variables on the graph. Using that graphical model
the system is then able to propagate and find the most likely configuration of the represented
variables, for instance the semantic category of a specific room is modelled as an hidden variable
with states representing the semantic values, by calculating probabilities on that variable the
system finds its belief on the room belonging to a specific category.

Since the graphical model produced by the conceptual map tries to model the distribution of
variables assuming the knowledge of the agent holds true, it represents the way to model
$P(x|known)$, and allows to calculate the density probability that the given set of sensed
features $x$ are sensed given that all the variables and graph structure, including the category of
a specific room $a$ are correctly modelled by the agent knowledge.
This way the distribution modelled by a factor graph equivalent to the chain-graph used by
the conceptual map can be used as an approximation for the conditional probability of $x$.

@@

With no knowledge on the unconditional probability the correct approach is to
assume a uniform distribution (\autoref{sec:max-entropy}).
That is represented in factor graphs by a graph without any factors: \autoref{fig:uniform-graph}.

@@

Thought, very often there is extra knowledge that can be obtained about the distribution
of the variables that helps to model the unconditional distribution. With that, more accurate
models can be produced.
In this case, the system is interested in detecting only if a specific variable category is not
known. With that in mind, it was assumed that an unknown category can only influence variables
by using the same structure as the known variables. This is: it was assumed the structure of the
graph was correct even when a room has an unknown category.

@@

If the structure and all the other variables can be assumed to be correctly modelled by the
knowledge of the agent, then all variables that were directly dependent on $a$ become directly
dependent between them, introducing a single factor (\autoref{fig:known-struct}).
By approximating this factor the agent can then obtain a model for the unconditional probability.

@@

\begin{figure}[h]
\centering
\subfloat[\label{fig:uniform-model}Uniform Model]{
\begin{tikzpicture}
  \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
    \& \& \node (room3) [latent] {}; \& \\
    \& \& \& \\
    \&
    \node (room1) [latent,draw=none] {}; \& \&
    \node (room2) [latent] {}; \\
    \& \& \& \\
    \&
    \&
    \&
    \node (factor2) [factor] {}; \\
    \node (shape1l) [obs] {$S_p$}; \&
    \node (appearance1l) [obs] {$A_p$}; \&
    \node (object1l) [obs] {$O_p$}; \&
    \node (prop2) [obs] {$X_p$}; \\
  };

  \draw [-] (room2) -- (factor2) -- (prop2);
  \draw [-] (room2) -- (room3) node (r23f) [midway,factor] {};

>>>>

%\subsection{Novelty Detection on a Variable Set of Features}
Note also that often novelty detection is applied on a fixed set of features
together with an assumption of a uniform unconditional probability.
On those cases $P(x)$ becomes a constant and therefore a novelty threshold
can be directly applied on $P(x|\overline{novel})$ as is the case in \cite{bishop1994novelty}.
But in the case where the set of features $x$ is variable it cannot be
discarded. There $P(x)$ plays a role in levering all the conditional
probabilities on different sets of variables into the same measure units
such that a threshold can be implemented.

\subsection{Assumption on a constant $P(novel)$}
The ratio between the conditional and unconditional probabilities of the sensed variables
is only directly applicable to novelty detection under the assumption of a constant $P(novel)$.

All the literature focus on novelty detection within a fixed scenario, where the set of variables
used to model the distribution is fixed and defined at the moment the threshold is trained.
Due to that, they can drop a constant $P(novel)$ and never need to directly or indirectly calculate
it.
To the best of the author knowledge there has not been studies on how to approximate it on dynamic
sets and structures of sensed variables, and so a strong assumption on being constant through all
the possible scenarios is used to allow dropping the factor.

This assumption was considered questionable on whether it holds on a realistic scenario,
and for that a sampling scheme able to comply with it is now presented as well an unbiased sampling
scheme. A method to generate samples that match this \emph{criterium} of having a constant $P(novel)$ can be:

\begin{algorithm}[Draw samples with equal $P(novel)$]
\begin{enumerate}
\item Sample a graph structure $G$ (where variable $a$ can be any of the known or unknown classes)
\item Decide novelty of $a$.
\item Sample the remaining variables according to distribution $P_G(x|a)$.
\end{enumerate}
\end{algorithm}

According to this sampling algorithm, given any structure the probability of
variable $a$ being novel is constant. This is somehow unexpected as it means
the sampling method is biased to adjust the real distribution $G$ such that
exactly a constant percentage of drawn samples are novel.

Alternative, the sampling method described below is considered more correct as
it introduces no bias on the distribution $G$ describing the real distribution
and classes of $a$:
\begin{algorithm}[Draw samples with equal $P(novel)$]
\begin{enumerate}
\item Sample a graph structure $G$ (where variable $a$ can be any of the known or unknown classes)
\item Sample variables $x$ according to distribution $P_G(x)$.
\end{enumerate}
\end{algorithm}

Due to the highly bias needed to drawn a constant amount of novel sampes from any structure, the
author defends this assumption to be very strong and points that future work should avoid.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Novelty Detection on the Conceptual Map}

@@

\subsection{Approximating the Conditional Probability}
The actual semantic mapping of the environment happens using the \emph{conceptual map}
introduced on \autoref{sec:conceptual-map}. During it the agent instantiates a
\emph{chain graph} to model distribution of the sensed variables according to the known
concepts and categories introduced as hidden variables on the graph. Using that graphical model
the system is able to propagate and find the most likely configuration of the represented
variables. For instance, the semantic category of a specific room is modelled as an hidden variable
with states representing the semantic values, by calculating probabilities on that variable the
system finds its belief on the room belonging to a specific category.

Since the graphical model produced by the conceptual map tries to model the distribution of
variables assuming the knowledge of the agent holds true, it represents a method to model
$P(x|known)$ since it allows to calculate the density probability that the set of features $x$ is
sensed given that all the variables and graph structure, including the category of
a specific room $a$ are correctly modelled by the agent knowledge.
This way the distribution modelled by a factor graph equivalent to the chain-graph used by
the conceptual map can be used as an approximation for the conditional probability of $x$.

@@

With no knowledge on the unconditional probability the correct approach is to
assume a uniform distribution (\autoref{sec:max-entropy}), which can be modelled by a
factor graph without any factors: \autoref{fig:uniform-graph}.

@@

Thought, very often there is extra knowledge that can be obtained about the distribution
of the variables that helps to model the unconditional distribution. With that, more accurate
models can be produced.
In this case, the system is interested in detecting only if a specific variable category is not
known. With that in mind, it was assumed that the graph structure is known and that an unknown
category can only influence variables by using the same structure as the known variables.
This way the structural information and all the other hidden variables the system is aware of
can be used to more accurately approximate the unconditional distribution.

@@

If the structure and all the other variables can be assumed to be correctly modelled by the
knowledge of the agent, then all variables that are directly dependent on $a$ become directly
dependent between them, introducing a single factor (\autoref{fig:known-struct}).
By approximating this factor the agent can then obtain a model for the unconditional probability.

@@

\begin{figure}[h]
\centering
\subfloat[\label{fig:uniform-model}Uniform Model]{
\begin{tikzpicture}
  \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
    \& \& \node (room3) [latent] {}; \& \\
    \& \& \& \\
    \&
    \node (room1) [factor] {}; \& \&
    \node (room2) [latent] {}; \\
    \& \& \& \\
    \&
    \&
    \&
    \node (factor2) [factor] {}; \\
    \node (shape1l) [obs] {$S_p$}; \&
    \node (appearance1l) [obs] {$A_p$}; \&
    \node (object1l) [obs] {$O_p$}; \&
    \node (prop2) [obs] {$X_p$}; \\
  };

  \draw [-,dashed] (room1) -- (shape1l);
  \draw [-,dashed] (room1) -- (appearance1l);
  \draw [-,dashed] (room1) -- (object1l);
  \draw [-] (room2) -- (prop2);

  \draw [-,dashed] (room1) -- (room2);
  \draw [-,dashed] (room1) -- (room3);
  \draw [-] (room2) -- (room3) node (r23f) [midway,factor] {};
  \node (captroom1) [above=0pt of room1] {\footnotesize{U}};
